Trajectories

### 2023-02-01 docs/index.qmd
Update manuscript

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        code-copy: hover
        reference-location: margin
        code-overflow: scroll
    pdf: default
link-external-newwindow: true
code-fold: true
code-line-numbers: true
fig-dpi: 1000
warning: false
echo: false
toc: true
colorlinks: true
number-sections: true
bibliography: "references.bib"
csl: "apa7.csl"
---

@@

# import packages
library(conflicted)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(purrr)
library(gt)
library(forcats)
library(stringr)
library(tibble)
library(tidybayes)
library(janitor)
library(brms)
library(scales)
library(marginaleffects)
library(lubridate)
library(bvqdev)

# load targets as R objects
tar_load(participants)
tar_load(df)
tar_load(items)
tar_load(bvq_data)
tar_load(model_fit_0)
tar_load(model_fit_1)
tar_load(model_fit_2)
tar_load(model_fit_3)
tar_load(model_fit_4)
tar_load(model_fit_4_prior)
tar_load(model_fit_prior)
tar_load(model_loos)
tar_load(rope_interval)
tar_load(posterior_description)
tar_load(posterior_description_re)

@@

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient)
```

@@

```{r}
#| label: tbl-items-summary
#| tbl-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
# items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
# 
# items %>%
#     unnest(list) %>%
#     group_by(language, list) %>%
#     summarise_at(
#         vars(freq_zipf, n_phon, lv),
#         list(
#             n = ~sum(!is.na(.)),
#             mean = ~mean(., na.rm = TRUE),
#             sd = ~sd(., na.rm = TRUE),
#             min = ~min(., na.rm = TRUE),
#             max = ~max(., na.rm = TRUE)
#         )
#     ) %>%
#     select(-c(n_phon_n, lv_n)) %>%
#     rename(n = freq_zipf_n) %>%
#     gt(groupname_col = "language", rowname_col = "list") %>%
#     fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>%
#     fmt_percent(matches("lv_"), decimals = 1) %>%
#     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>%
#     tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>%
#     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
#     cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>%
#     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
#     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
#     cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>%
#     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
#     cols_merge_range(col_begin = lv_min, col_end = lv_max) %>%
#     cols_label(
#         list = "",
#         n = md("*N*"),
#         freq_zipf_mean = md("Mean ± SD"),
#         freq_zipf_min = md("Range"),
#         n_phon_mean = md("Mean ± SD"),
#         n_phon_min = md("Range"),
#         lv_mean = md("Mean ± SD"),
#         lv_min = md("Range")
#     ) %>%
#     tab_style(
#         cell_text(style = "italic"),
#         cells_column_labels(everything())
#     ) %>%
#     tab_style(
#         cell_text(weight = "bold"),
#         cells_column_spanners(everything())
#     ) %>%
#     summary_rows(
#         fns = list("Grand mean" = ~mean(.)),
#         columns = ends_with("_mean")
#     ) %>%
#     summary_rows(
#         fns = list("N" = ~sum(.)),
#         columns = n,
#         formatter = fmt_integer
#     )
```

@@

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = freq_zipf)) +
    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: fig-items-frequency-phonemes
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq_zipf) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq_zipf) %>% 
    ggplot(aes(n_phon, freq_zipf)) +
    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
               position = position_jitter(width = 0.1), seed = 888) +
    geom_smooth(method = "lm", formula = "y ~ x" ,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq_zipf) %>% 
    count(freq_zipf, lv) %>% 
    ggplot(aes(freq_zipf, lv, size = n)) +
    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq_zipf) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):

@@

```{r}
#| label: fig-exposure-distribution
#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
#| warnings: false
#| fig-height: 7
#| fig-width: 13
df %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        )
    ) %>%
    count(exposure, language) %>%
    ggplot(aes(exposure, n, fill = language)) +
    facet_wrap( ~ language) +
    geom_col(colour = "white") +
    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
         y = "Number of observations",
         fill = "Language") +

    df %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        ),
        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
    ) %>%
    count(exposure, freq, language) %>%
    ggplot(aes(
        x = freq,
        y = exposure,
        colour = language,
        size = n
    )) +
    facet_wrap( ~ language) +
    geom_point() +
    labs(
        x = "Frequency (Zipf score)",
        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
        colour = "Language",
        fill = "Language",
        size = "Number of observations"
    ) +
    theme(panel.grid = element_blank()) +
    plot_layout(ncol = 1) &
    scale_colour_manual(values = clrs[c(1, 4)]) &
    scale_fill_manual(values = clrs[c(1, 4)]) &
    guides(fill = "none", colour = "none") &
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank())
```

@@

```{r}
#| label: fig-likelihood
#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire. Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution. For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*. *Familiarity* describes a given participant's familiarity with a given word. When familiarity is low (below some threshold 1), the reulting response is *No*. When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension). When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production). These thresholds are estimated in the cumulative regression model as intercepts 1, and 2. The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely. See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
#| fig-width: 7
#| fig-height: 3.5
dist_df = tribble(~mean, ~sd, 0,   1)

responses <- data.frame(
    ymin = c(-Inf, -1.5, 1.5),
    ymax = c(-1.5, 1.5, Inf),
    xmin = -Inf,
    xmax = Inf,
    .category = str_wrap(c("No", "Understands", "Understands and Says"), width = 12)
)

responses %>%
    ggplot(aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = .category)) +
    geom_rect(alpha = 0.5) +
    stat_dist_slab(
        data = dist_df, 
        aes(dist = "norm", arg1 = mean, arg2 = sd),
        position = "dodge", inherit.aes = FALSE,
        fill = NA, colour = "black"
    ) +
    geom_hline(yintercept = c(-1.5, 1.5), linetype = "dotted", size = 1) +
    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
              vjust = 1, hjust = 0.5) +
    annotate(
        geom = "text",
        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
        y = c(-1.5, 2), parse = TRUE,
        x = 0.5, angle = 90, vjust = -0.5
    ) +
    labs(y = "Hypothetical response-generating latent distribution",
         x = "Response probability",
         fill = "Response type",
         colour = "Response type") +
    coord_flip() +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_x_continuous(labels = percent) +
    scale_y_continuous(limits = c(-4, 4)) +
    theme(panel.grid = element_blank(),
          legend.position = "none")
```

@@

```{r}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(
        values = rev(clrs), 
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: fig-prior-predictions
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

```{r}
#| lab: fig-responses-age-phonemes
#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
df %>% 
    mutate(
        age = floor(age),
        n_phon = cut_interval(n_phon, n = 4)
    ) %>%
    count(age, n_phon, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~n_phon) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@

#### Correlations

>>>>

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        code-copy: hover
        reference-location: margin
        code-overflow: scroll
link-external-newwindow: true
code-fold: true
code-line-numbers: true
fig-dpi: 1000
page-layout: full
warning: false
echo: false
toc: true
colorlinks: true
number-sections: true
bibliography: "references.bib"
csl: "apa7.csl"
---

@@

# import packages
library(conflicted)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(purrr)
library(gt)
library(forcats)
library(stringr)
library(tibble)
library(tidybayes)
library(janitor)
library(glue)
library(brms)
library(scales)
library(marginaleffects)
library(lubridate)
library(bvqdev)

# load targets as R objects
tar_load(participants)
tar_load(responses)
tar_load(items)
tar_load(bvq_data)
tar_load(model_fit_0)
tar_load(model_fit_1)
tar_load(model_fit_2)
tar_load(model_fit_3)
tar_load(model_fit_4)
tar_load(model_fit_4_prior)
tar_load(model_fit_prior)
tar_load(model_loos)
tar_load(posterior_draws)
tar_load(posterior_draws_re)

@@

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient)

rope_interval <- c(lower = -0.1, upper = +0.1)
```

@@

```{r}
#| label: tbl-items-summary
#| tbl-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
# items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
# 
# items %>%
#     unnest(list) %>%
#     group_by(language, list) %>%
#     summarise_at(
#         vars(freq, n_phon, lv),
#         list(
#             n = ~sum(!is.na(.)),
#             mean = ~mean(., na.rm = TRUE),
#             sd = ~sd(., na.rm = TRUE),
#             min = ~min(., na.rm = TRUE),
#             max = ~max(., na.rm = TRUE)
#         )
#     ) %>%
#     select(-c(n_phon_n, lv_n)) %>%
#     rename(n = freq_n) %>%
#     gt(groupname_col = "language", rowname_col = "list") %>%
#     fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>%
#     fmt_percent(matches("lv_"), decimals = 1) %>%
#     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>%
#     tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>%
#     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
#     cols_merge_uncert(col_val = freq_mean, col_uncert = freq_sd) %>%
#     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
#     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
#     cols_merge_range(col_begin = freq_min, col_end = freq_max) %>%
#     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
#     cols_merge_range(col_begin = lv_min, col_end = lv_max) %>%
#     cols_label(
#         list = "",
#         n = md("*N*"),
#         freq_mean = md("Mean ± SD"),
#         freq_min = md("Range"),
#         n_phon_mean = md("Mean ± SD"),
#         n_phon_min = md("Range"),
#         lv_mean = md("Mean ± SD"),
#         lv_min = md("Range")
#     ) %>%
#     tab_style(
#         cell_text(style = "italic"),
#         cells_column_labels(everything())
#     ) %>%
#     tab_style(
#         cell_text(weight = "bold"),
#         cells_column_spanners(everything())
#     ) %>%
#     summary_rows(
#         fns = list("Grand mean" = ~mean(.)),
#         columns = ends_with("_mean")
#     ) %>%
#     summary_rows(
#         fns = list("N" = ~sum(.)),
#         columns = n,
#         formatter = fmt_integer
#     )
```

@@

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = freq)) +
    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: fig-items-frequency-phonemes
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
               position = position_jitter(width = 0.1), seed = 888) +
    geom_smooth(method = "lm", formula = "y ~ x" ,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv, size = n)) +
    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):

@@

```{r}
#| label: fig-exposure-distribution
#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
#| warnings: false
#| fig-height: 7
#| fig-width: 13
responses %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        )
    ) %>%
    count(exposure, language) %>%
    ggplot(aes(exposure, n, fill = language)) +
    facet_wrap( ~ language) +
    geom_col(colour = "white") +
    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
         y = "Number of observations",
         fill = "Language") +

    responses %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        ),
        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
    ) %>%
    count(exposure, freq, language) %>%
    ggplot(aes(
        x = freq,
        y = exposure,
        colour = language,
        size = n
    )) +
    facet_wrap( ~ language) +
    geom_point() +
    labs(
        x = "Frequency (Zipf score)",
        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
        colour = "Language",
        fill = "Language",
        size = "Number of observations"
    ) +
    theme(panel.grid = element_blank()) +
    plot_layout(ncol = 1) &
    scale_colour_manual(values = clrs[c(1, 4)]) &
    scale_fill_manual(values = clrs[c(1, 4)]) &
    guides(fill = "none", colour = "none") &
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank())
```

@@

```{r}
#| label: fig-likelihood
#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire. Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution. For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*. *Familiarity* describes a given participant's familiarity with a given word. When familiarity is low (below some threshold 1), the reulting response is *No*. When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension). When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production). These thresholds are estimated in the cumulative regression model as intercepts 1, and 2. The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely. See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
#| fig-width: 7
#| fig-height: 3.5
dist_df <- tribble(~mean, ~sd, 0,   1)

thresholds <- data.frame(
    ymin = c(-Inf, -1.5, 1.5),
    ymax = c(-1.5, 1.5, Inf),
    xmin = -Inf,
    xmax = Inf,
    .category = str_wrap(c(
        "No", 
        "Understands",
        "Understands and Says"
    ),
    width = 12)
)

thresholds %>%
    ggplot(aes(ymin = ymin, ymax = ymax,
               xmin = xmin, xmax = xmax,
               fill = .category)) +
    geom_rect(alpha = 0.5) +
    stat_dist_slab(
        data = dist_df, 
        aes(dist = "norm", arg1 = mean, arg2 = sd),
        position = "dodge", inherit.aes = FALSE,
        fill = NA, colour = "black"
    ) +
    geom_hline(yintercept = c(-1.5, 1.5), 
               linetype = "dotted",
               linewidth = 1) +
    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
              vjust = 1, hjust = 0.5) +
    annotate(
        geom = "text",
        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
        y = c(-1.5, 2), parse = TRUE,
        x = 0.5, angle = 90, vjust = -0.5
    ) +
    labs(y = "Hypothetical response-generating latent distribution",
         x = "Response probability",
         fill = "Response type",
         colour = "Response type") +
    coord_flip() +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_x_continuous(labels = percent) +
    scale_y_continuous(limits = c(-4, 4)) +
    theme(panel.grid = element_blank(),
          legend.position = "none")
```

@@

```{r}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(
        values = rev(clrs), 
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: fig-prior-predictions
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

```{r}
#| lab: fig-responses-age-phonemes
#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
responses %>% 
    mutate(
        age = floor(age),
        n_phon = cut_interval(n_phon, n = 4)
    ) %>%
    count(age, n_phon, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~n_phon) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@
